\vspace{-1cm} Model linearity breeds contempt: using Bayesian non-linear models to uncover broad macroecological patterns

### 2021-05-01 manuscript/manuscript.tex
Multithread attempt and manuscript
Some changes t plots and some changes to the processing of the data
Sliding data and code

%Title paper
\title{\vspace{-1cm}
Model linearity breeds contempt: using Bayesian non-linear models to uncover general biogeographical patterns}
% Model familiarity breeds contempt: simple models to compare properties of species' distributions
% Model simplicity breeds contempt: simple models to compare properties of species' distributions
% Model complexity breeds contempt: using non-linear models to compare basic properties of species' distributions
\author[1,*]{\normalsize Bernat Bramon Mora}
\author[1]{\normalsize Jake M.\ Alexander}
\affil[1]{\footnotesize Institute of Integrative Biology, ETH Zürich, Zürich, Switzerland}
\affil[*]{\footnotesize  bernat.bramon@gmail.com}

@@

%Consider testing tails of distribution? I would use the Cauchy or generalized normal distribution for heavy tail distribution without skewness; and I would use an exponentially modified Gaussian distribution with heavy tails for a skew distribution with heavy tails. 
% Option (1) for a long-tail distribution: Generalized normal distribution. We model it such that shape parameter beta can go from 1 to 8
% Option (2) for a long-tail distribution: Student's t-distribution. In this case, we would need to define the shape parameter so that it is larger than 2 (otherwise, we would have undefined variance)
% Option (1) for a long-tail skewed distribution. The exponentially modified Gaussian distribution has potential, but I am not sure I can fit a parameter that goes from like lambda if I want the skewness in either side.
% Option (2) for a long-tail skewed distribution. Version 2 of the Generalized normal distribution. The problem here is that this function has a few discontinuities, which means that one needs to use a step function. I believe this to be quite problematic in Stan...
% Option (3) for a long-tail skewed distribution. Skewed student-t distribution. This seems to be complex but quite what I need. It will be hard to fit but seems quite promising. See:
%https://www.tandfonline.com/doi/abs/10.1080/01621459.1998.10474117
%https://hess.copernicus.org/preprints/hess-2015-24/
%https://agupubs.onlinelibrary.wiley.com/doi/epdf/10.1029/2009WR008933

@@

The value of gaining a mechanistic understanding of species' distributions is unquestionable (ref), with several studies highlighting the importance of factors such as biotic interactions and dispersal ability in setting species' range limits \citep{wiszRoleBioticInteractions2013, pollockUnderstandingCooccurrenceModelling2014, neuschulzBioticInteractionsSeed2018}. That said, a lot can be learned from taking a phenomenological approach, focussing instead on the description of basic properties of species' realized niches. For example, the study of species' range sizes along environmental gradients can reveal general biodiversity patterns that are crucial from a conservation and management perspective \citep{stevensElevationalGradientAltitudinal1992}. Differences in species' responses to the environment could shed light on how climatic processes and historical contingencies have shaped their distributions \citep{rohdeLatitudinalGradientsSpecies1992, more in Rapoport}. Other properties, such as the skewness of species' distributions, can also reveal general underlying processes regarding species' physiological tolerance to different environmental conditions \citep{kaufmanDiversityNewWorld1995}. More generally, understanding the shape of species' realized niches and the extend to which these vary across species is a crucial issue in ecology and biogeography (ref); however, we do not have an effective way to parsimoniously compare the realized niches of many species. Indeed, there is no general agreement on the shape of species' distributions (ref).

@@

\subsubsection*{Skewed response curve}
When species experience abiotic or biotic pressures that increase or decrease along an environmental gradient, one might expect their distributions to be skewed in one direction. Likewise, this same skewed shape can also emerge as a result of species' asymmetric environmental tolerance. One way to accommodate this feature to our models is by considering a skewed normal distribution (Supplementary Figure XX). As for the two cases described above, the Gaussian is a special case of this distribution, and it contains a parameter that controls for the level and direction of `skewness' (Fig~\ref{fig:response}c). Importantly, this distribution presents normal-like tails; therefore, the added skewness does not make additional assumptions regarding how species are distributed along the gradient. To test for the existence of this feature, we modified Eq.~(\ref{eq:baseline-response}) as
\begin{equation}
\begin{split}
p_{ij} & =\hat{p}_{ij} \left[1+\text{erf}\left(\lambda_{i}\left(x_{j}-\beta^{\prime}_{i}\right)\sqrt{\frac{\gamma^{\prime}_{i}}{2}}\right)\right] \\
\text{log}\left(\hat{p}_{ij}\right) & = -\alpha^{\prime}_{i} - \gamma^{\prime}_{i} \left(x_{j}-\beta^{\prime}_{i}\right)^2 ,
\end{split}
\label{eq:skewed-response}
\end{equation}
where $\alpha^{\prime}_{i} = q_{1}\left(\alpha_{i}, \gamma^{\prime}_{i}, \lambda_{i}\right)$, $\beta^{\prime}_{i} = q_{2}\left(\gamma^{\prime}_{i}, \lambda_{i}\right)$, $\gamma^{\prime}_{i} = q_{3}\left(\gamma_{i}, \lambda_{i}\right)$, and $\lambda_{i}$ is a parameter that describes the skewness of the distribution, which we define as $\lambda_{i}\in\left(-10, 10\right)$. Notice that these boundary conditions for $\lambda_{i}$ are arbitrarily set for computational purposes, as this parameter is theoretically not bounded. The function `erf' is the error function (see \citealt{ashourApproximateSkewNormal2010}). We choose an adaptive prior such that $\lambda_{i}\sim \text{Normal}\left(\hat{\lambda}, \sigma_{\lambda}\right)$, where $\hat{\lambda}\sim\text{Normal}\left(0, 1\right)$ and $\sigma_{\lambda}\sim\text{Exponential}\left(1\right)$. This model can be re-parametrized following $q_1$, $q_2$ and $q_3$, and set the rest of the prior distributions as described for the baseline model (see Supplementary Table 2; Code Availability section). In this case, the Gaussian distribution is a special case of Eq.~(\ref{eq:skewed-response}) when $\lambda_{i}=0$ \citep{ashourApproximateSkewNormal2010}.

One way to test the extend to which species' physiological traits inform the skewness of the distribution of any species $i$ is by characterizing $\lambda_{i}$ as a Gaussian process. That is, we can consider the prior distribution for $\lambda_{i}$ as a multivariate normal with a variance covariance matrix $\Sigma^{\lambda}$. As described before, this variance-covariance matrix can be estimated following Eq.~(\ref{eq:covariance-baseline}), and its structure can shed light on how informative the physiological traits are. Likewise, we could instead use Eq.~(\ref{eq:covariance-complex}) and test the effectiveness of other prior information in explaining the skewness of species' distributions.

\subsubsection*{Heavy-tail and skewed response curve}
Finally, one could consider a response curve with both kurtosis and skewness. A convenient way to achieve this is by using a response curve that follows a skewed generalized error distribution (also see XX). This is a combination of the two distributions described above, containing two parameters that control for both the level and direction of kurtosis and skewness (Fig~\ref{fig:response}d). The skewed generalized error distribution can be considered by modifying the species' response curve in Eq.~(\ref{eq:baseline-response}) as
\begin{equation}	
\begin{split}
\text{log}\left(p_{ij}\right) = -\alpha_{i} - \left(\frac{\gamma^{\prime}_{i}\,|x_{j}-\beta^{\prime}_{i}|}{1+\lambda_{i}\,\text{sgn}\left(x_{j}-\beta^{\prime}_{i}\right)}\right)^{\nu_{i}} ,
\end{split}
\label{eq:heavy-skewed-response}
\end{equation}
where $\gamma^{\prime}_{i} = f_{1}\left(\gamma_{i}, \nu_{i}, \lambda_{i}\right)$, $\beta^{\prime}_{i} = f_{2}\left(\gamma_{i}, \beta_{i}, \nu_{i}, \lambda_{i}\right)$, $\nu_{i}$ and $\lambda_i$ are parameters that describe the kurtosis and skewness of the distribution, respectively. The function $\text{sgn}\left(x\right)$ characterizes the sign function. We define $\nu_i$ and its prior as in Eq.~\ref{eq:generalized-response}. However, $\lambda_i$ is now defined as $\lambda_{i}\in\left(-1, 1\right)$, with an adaptive prior such that $\text{logit}\left(\frac{\lambda_{i}+1}{2}\right)\sim \text{Normal}\left(\hat{\lambda}, \sigma_{\lambda}\right)$, where $\hat{\lambda}\sim\text{Normal}\left(0, 1\right)$ and $\sigma_{\lambda}\sim\text{Exponential}\left(1\right)$. Again, we can re-parametrize the model following $f_1$ and $f_2$, and set the rest of the prior distributions as in the baseline model (see Supplementary Table 2; Code Availability section). Notice that the generalized error distribution (Eq.~\ref{eq:generalized-response}) and the skew normal distribution (Eq.~\ref{eq:skewed-response}) are special cases of Eq.~(\ref{eq:heavy-skewed-response}) when $\lambda_i=0$ and $\nu_i=2$, respectively.

\section*{Results}
We studied the distribution data to characterize species' realized niches. To do so, we selected the main axis of variation of all environmental variables (Supplementary Fig.~XX). Then, using the presence and absence of species across sites as the response variable, we sampled the posterior distributions of the baseline model, accounting for the information in the floristic database regarding species' indicator values and range of variation. Studying the relationship between the mean and variance of the distributions, we found clear evidence of the Rapopor's rule. Specifically, we found that species' range sizes decline with elevation and temperature (i.e.~$\beta_{i}$ and $\gamma_{i}$ in the baseline model are negatively correlated; Fig.~\ref{fig:correlation}). The relationship was also found when using instead ordinal data (Supplementary Fig.~XX), and it does not seem to depend on species' type or reflect species' abundance change tendency over the years (Fig.~\ref{fig:universality}). This pattern was not present along the second axis of variation of the environmental variables (Supplementary Fig.~XX). The comparison between the other parameter estimates revealed additional, somewhat more expected, relationships. In particular, we found the amplitude of distributions to be positively and negatively correlated with the mean and the variance along the elevation gradient, respectively (Supplementary Fig.~XX).

@@

\begin{figure}[ht]
  \centering
    \includegraphics[width=0.9\textwidth]{figures/means-Rapopor}
    	  \vspace{0.1cm}
	   \caption{Universality of the relationship between mean and variance of species' distributions. Comparison between how different types of species are mapped in Fig.~\ref{fig:correlation}a. Panel (a) describes the correlation coefficient between $\beta_i$ and $\gamma_i$ for each type of species. Panel (b) shows the differences between neophytes and native species in the way these are distributed along the environmental gradient. Panel (c) shows the same differences for species that have decreased, decreased in low elevations, increase and remain stable over the last decades (see Supplementary Table 1 for further details). {\color{red} I might move this to the Supplementary Information.}}
	   %7.5-5.5
      \label{fig:universality}
\end{figure}

Maintaining the symmetry of species' distributions, we then allowed the kurtosis---or shape of the tails---of these to vary in different ways. To do so, we changed the response curve of our Bayesian model to follow a generalized error distribution (Eq.~\ref{eq:generalized-response}). A comparison of the WAIC values showed this non-linear regression to outperform the baseline model (Supplementary Fig.~XX)\footnote[2]{\label{note1} I am still waiting on the results for this (currently running in the cluster), and this is just my prior expectation based on what I've seen in some of the other models I've been working with.}. Studying the resulting posterior distributions, we found the parameter controlling for the kurtosis to be centred around $\nu_i \sim 2$, which corresponds to a distribution with a Gaussian shape (Fig.~\ref{fig:kurtosis-skewness}). However, such parameter estimates displayed a lot of variation across species, which might indicate that the shape of the tails of the distribution is species-specific (Supplementary Fig.~XX).

\begin{figure}[ht]
  \centering
    \includegraphics[width=0.8\textwidth]{figures/kurto-skew}
    	  \vspace{0.1cm}
	   \caption{Average kurtosis and skewness of species' distributions. Calculated using the posterior distributions of parameters $\hat{\nu}$ and $\hat{\lambda}$ from the models (see Supplementary Table XX), the two panels describe the average (a) kurtosis and (b) skewness of distributions, respectively. Panel (a) displays the results obtained by using a response curve that follows a generalized error distribution. Panel (b) displays the results obtained by using a response curve that follows a skewed normal distribution. In both case, the red dotted line indicates the conditions by which species are normally distributed along the environmental axis.}
      \label{fig:kurtosis-skewness}
\end{figure}

Using Eq.~(\ref{eq:skewed-response}), we next studied the skewness of species distributions. Based on the estimates for the WAIC values, this model outperformed the rest (Supplementary Fig.~XX), which sheds light on the naturally skewed nature of species' distributions. Perhaps most importantly, studying the mean value of the skewness across species (i.e.~parameter $\hat{\lambda}$), we found that species' distributions present steeper declines towards lower elevations (Fig.~\ref{fig:kurtosis-skewness}). Moreover, species' physiology seems to strongly shape this parameter, which suggest that distribution skewness is an intrinsic property of species (New Fig.~XX)\footref{note1}.

Finally, we used a model that allowed for heavy-tails and skewed response curves (Eq.~\ref{eq:heavy-skewed-response}) to understand the relationship between the kurtosis and skewness of distributions. This model outperformed the rest, presenting Akaike weights close to 1 when comparing models (Supplementary Fig.~XX). Comparing the posterior distribution for parameters $\nu_i$ and $\lambda_i$, we found this to be uncorrelated\footref{note1}. However, different types of species seem to present characteristically different distributions based on their kurtosis and skewness. Importantly, this model identified similar relationships between parameters as the baseline model (Supplementary Fig.~XX), and measured similar average levels of kurtosis and skewness than the the other two models (Fig.~\ref{fig:kurtosis-skewness})\footref{note1}.

@@

>>>>

%Title paper
\title{\vspace{-1cm}
Model linearity breeds contempt: using Bayesian non-linear models to uncover broad macroecological patterns}
% Model familiarity breeds contempt: simple models to compare properties of species' distributions
% Model simplicity breeds contempt: simple models to compare properties of species' distributions
% Model complexity breeds contempt: using non-linear models to compare basic properties of species' distributions
\author[1,*]{\normalsize Bernat Bramon Mora}
\author[1]{\normalsize Jake M.\ Alexander}
\affil[1]{\footnotesize Institute of Integrative Biology, ETH Zürich, Zürich, Switzerland}
\affil[*]{\footnotesize  bernat.bramon@gmail.com}

@@

% REFERENCES TO LOOK AT FROM JAKE
% DATE EMAIL: 16.03.2021

%https://www.tandfonline.com/doi/abs/10.1080/01621459.1998.10474117
%https://hess.copernicus.org/preprints/hess-2015-24/
%https://agupubs.onlinelibrary.wiley.com/doi/epdf/10.1029/2009WR008933

%############# NEW INTRO ##############

% Describe different aspects of shape

One of the central goals of ecology is to understand the ways species are distributed across space and time (ref). Over the last two decades, ecologists have developed a plethora of distribution models to try to untangle the factors that play a role in defining such distributions \citep{guisanPredictiveHabitatDistribution2000}. The idea is that by untangling the ways in which such biotic and abiotic factors shape species' distributions, we can gain a mechanistic understanding on how ecological communities are established and change over time. While these models are fundamental to the scientific community for predicting changes in species geographic distributions and the effects of environmental disturbances, distribution models ... can increase the predictive performance of some of the models \citep{norbergComprehensiveEvaluationPredictive2019},

While the value of these models is unquestionable, these approaches often rely on broad simplifications regarding the shape of species' realized niches. Describing

 These models estimate species' realized niches using several covariates, including environmental variables \citep{guisanPredictingSpeciesDistribution2005}, species ecological traits' \citep{pollockRoleFunctionalTraits2012} and phylogenetic relations \citep{ivesGeneralizedLinearMixed2011}. More recently, some of the focus have shifted towards approaches that estimate and account for biotic factors, such as competitive or facilitative relationships between species \citep{ovaskainenHowMakeMore2017}. The idea is that by untangling the ways in which such biotic and abiotic factors shape species' distributions, we can gain a mechanistic understanding on how ecological communities are established and change over time. However, while these factors can increase the predictive performance of some of the models \citep{norbergComprehensiveEvaluationPredictive2019}, the interpretation of the corresponding parameter estimates has been recently questioned \citep{harrisInferringSpeciesInteractions2016, thurmanTestingLinkSpecies2019, poggiatoInterpretationsJointModeling2021}. This was best illustrated by \citet{blanchetCooccurrenceNotEvidence2020}, who used basic statistical arguments to highlight the artefactual nature of the link between co-occurrence and species' ecological interactions drawn by some distribution models.

From a phenomenological perspective, one can focus on several aspects of species' distributions. For example,

Two aspects that are far less explored: the shape and the relationship between the properties of species distributions.

Many ecological textbooks \citep{krebsEcologyExperimentalAnalysis1972} assume the shape of species distributions to be unimodal and symmetric. However, some have warned that empirical distributions can take many different forms \citep{austinModelsAnalysisSpecies1987, austinSpatialPredictionSpecies2002}. From a purely statistical perspective, there is a strong argument to be made in support of bell shaped distributions. That is, if all that we are willing to assume about a species' distributions is that these occupy finite geographic ranges---i.e.~their probability distributions have finite variance---the most conservative statistical approach is to model their distribution as Gaussian \citep{frankCommonPatternsNature2009}. That said, many natural processes can lead to non-normal distributions, and heavy-tails and skewed shapes are very common across fields. For example, heavy-tailed distributions can capture processes such as seasonality (e.g.~in communications patterns; \citealt{malmgrenPoissonianExplanationHeavy2008}) or some stochastic events (e.g.~in the spread of infectious diseases; \citealt{wongEvidenceThatCoronavirus2020}). Indeed, species' dispersal patterns have been shown to have heavy-tails due to natural variability among individuals \citep{petrovskiiDispersalStatisticallyStructured2009}. Likewise, several processes can lead to skewed distributions. For example, when species experience abiotic or biotic pressures that increase or decrease along an environmental gradient, one might expect their distributions to be skewed in one direction; and this same skewed shape can also emerge as a result of species' asymmetric environmental tolerance.

 For example, species' distributions are shaped by multiple stochastic events, such as di distributions to showcase heavy-tails or a skewed shapes, revealing interesting ecological processes shaping biodiversity patterns \citep{austinNonlinearSpeciesResponse1976, minchinEvaluationRelativeRobustness1987}.

% Relate them to macroecological hypothesis

That said, a lot can be learned about species' geographic distributions from taking a phenomenological approach, focussing instead on the description of basic properties of their realized niches. For example, the study of species' range sizes along environmental gradients can reveal general biodiversity patterns that are crucial from a conservation and management perspective \citep{stevensElevationalGradientAltitudinal1992}. Differences in species' responses to the environment could shed light on how climatic processes and historical contingencies have shaped their distributions \citep{rohdeLatitudinalGradientsSpecies1992, more in Rapoport}. Other properties, such as the skewness of species' distributions, can also reveal general underlying processes regarding species' physiological tolerance to different environmental conditions \citep{kaufmanDiversityNewWorld1995}. More generally, understanding the shape of species' realized niches and the extend to which these vary across species is a crucial issue in ecology and biogeography (ref); however, we do not have an effective way to parsimoniously compare the realized niches of many species. Indeed, there is no general agreement on the shape of species' distributions (ref).

% Describe limitations in previous methods to characterize shape of Realized Niche: prior knowledge; simultaneous estimation of many (also rare); uncertainty

% Intro new method
Here, we instead rethink traditional modelling approaches and develop a conceptually simple---and yet statistical and computationally complex---statistical framework to revisit some classic hypothesis in ecology and biogeoraphy. In particular, we develop a Bayesian hierarchical model that accounts for all prior information that we have regarding the distribution of alpine plant species along an elevation gradient in the Swiss Alps, including expert knowledge on species environmental indicator values, range sizes, and plant physiology. We start by considering species' response curves as Gaussian distributed, and then we adapt our model to allow for skewed and long-tailed distributions. Using this statistical framework, we are able to compare the basic properties of the realized niches of multiple species, testing for the existence of general biogeographical patterns.

% Questions

%############# OLD INTRO ##############

@@

The value of gaining a mechanistic understanding of species' distributions is unquestionable (ref), with several studies highlighting the importance of factors such as biotic interactions and dispersal ability in setting species' range limits \citep{wiszRoleBioticInteractions2013, pollockUnderstandingCooccurrenceModelling2014, neuschulzBioticInteractionsSeed2018}. That said, a lot can be learned about species' geographic distributions from taking a phenomenological approach, focussing instead on the description of basic properties of their realized niches. For example, the study of species' range sizes along environmental gradients can reveal general biodiversity patterns that are crucial from a conservation and management perspective \citep{stevensElevationalGradientAltitudinal1992}. Differences in species' responses to the environment could shed light on how climatic processes and historical contingencies have shaped their distributions \citep{rohdeLatitudinalGradientsSpecies1992, more in Rapoport}. Other properties, such as the skewness of species' distributions, can also reveal general underlying processes regarding species' physiological tolerance to different environmental conditions \citep{kaufmanDiversityNewWorld1995}. More generally, understanding the shape of species' realized niches and the extend to which these vary across species is a crucial issue in ecology and biogeography (ref); however, we do not have an effective way to parsimoniously compare the realized niches of many species. Indeed, there is no general agreement on the shape of species' distributions (ref).

@@

\subsubsection*{Skewed response curve}
When species experience abiotic or biotic pressures that increase or decrease along an environmental gradient, one might expect their distributions to be skewed in one direction. Likewise, this same skewed shape can also emerge as a result of species' asymmetric environmental tolerance. One way to accommodate this feature to our models is by considering a skewed normal distribution (Supplementary Figure XX). As for the two cases described above, the Gaussian is a special case of this distribution, and it contains a parameter that controls for the level and direction of `skewness' (Fig~\ref{fig:response}c). Importantly, this distribution presents normal-like tails; therefore, the added skewness does not make additional assumptions regarding how species are distributed along the gradient. To test for the existence of this feature, we modified Eq.~(\ref{eq:baseline-response}) as
\begin{equation}	
\begin{split}
\text{log}\left(p_{ij}\right) = -\alpha_{i} - \gamma^{\prime}_{i}\left(\frac{x_{j}-\beta^{\prime}_{i}}{1+\lambda_{i}\,\text{sgn}\left(x_{j}-\beta^{\prime}_{i}\right)}\right)^{2} ,
\end{split}
\label{eq:skewed-response}
\end{equation}
where $\gamma^{\prime}_{i} = q_{1}\left(\gamma_{i}, \nu_{i}, \lambda_{i}\right)$, $\beta^{\prime}_{i} = q_{2}\left(\gamma_{i}, \beta_{i}, \nu_{i}, \lambda_{i}\right)$, and $\lambda_{i}$ is a parameter that describes the skewness of the distribution such that $\lambda_{i}\in\left(-1, 1\right)$. The function $\text{sgn}\left(x\right)$ characterizes the sign function.  We chose $\lambda_i$ to have an adaptive prior such that $\text{logit}\left(\frac{\lambda_{i}+1}{2}\right)\sim \text{Normal}\left(\hat{\lambda}, \sigma_{\lambda}\right)$, where $\hat{\lambda}\sim\text{Normal}\left(0, 1\right)$ and $\sigma_{\lambda}\sim\text{Exponential}\left(1\right)$. Notice that this model can be re-parametrized following $q_1$ and $q_2$, allowing us to set the rest of the prior distributions as described for the baseline model (see Supplementary Table 2; Code Availability section). In this case, the Gaussian distribution is a special case of Eq.~(\ref{eq:skewed-response}) when $\lambda_{i}=0$ \citep{ashourApproximateSkewNormal2010}.

\subsubsection*{Heavy-tail and skewed response curve}
Finally, one could consider a response curve with both kurtosis and skewness. A convenient way to achieve this is by using a response curve that follows a skewed generalized error distribution (also see XX). This is a combination of the two distributions described above, containing two parameters that control for both the level and direction of kurtosis and skewness (Fig~\ref{fig:response}d). The skewed generalized error distribution can be considered by modifying the species' response curve in Eq.~(\ref{eq:baseline-response}) as
\begin{equation}	
\begin{split}
\text{log}\left(p_{ij}\right) = -\alpha_{i} - \left(\frac{\gamma^{\prime}_{i}\,|x_{j}-\beta^{\prime}_{i}|}{1+\lambda_{i}\,\text{sgn}\left(x_{j}-\beta^{\prime}_{i}\right)}\right)^{\nu_{i}} ,
\end{split}
\label{eq:heavy-skewed-response}
\end{equation}
where $\gamma^{\prime}_{i} = f_{1}\left(\gamma_{i}, \nu_{i}, \lambda_{i}\right)$, $\beta^{\prime}_{i} = f_{2}\left(\gamma_{i}, \beta_{i}, \nu_{i}, \lambda_{i}\right)$, $\nu_{i}$ and $\lambda_i$ are parameters that describe the kurtosis and skewness of the distribution, respectively. We define $\nu_i$, $\lambda_i$ and their prior distributions as in Eq.~\ref{eq:generalized-response} and \ref{eq:skewed-response}, respectively. Again, we can re-parametrize the model following $f_1$ and $f_2$, and set the rest of the prior distributions as in the baseline model (see Supplementary Table 2; Code Availability section). Notice that the generalized error distribution (Eq.~\ref{eq:generalized-response}) and the skew normal distribution (Eq.~\ref{eq:skewed-response}) are special cases of Eq.~(\ref{eq:heavy-skewed-response}) when $\lambda_i=0$ and $\nu_i=2$, respectively.

One way to test the extend to which species' physiological traits inform the kurtosis or skewness of the distribution of any species $i$ is by characterizing $\nu_{i}$ or $\lambda_{i}$ as Gaussian processes. That is, we can consider the prior distribution for these parameters as multivariate normal with a variance covariance matrix $\Sigma$. As described before, this variance-covariance matrix can be estimated following Eq.~(\ref{eq:covariance-baseline}), and its structure can shed light on how informative the physiological traits are. Likewise, we could instead use Eq.~(\ref{eq:covariance-complex}) and test the effectiveness of other prior information in explaining the skewness of species' distributions.

\section*{Results}
We studied the distribution data to characterize species' realized niches. To do so, we selected the main axis of variation of all environmental variables (Supplementary Fig.~XX). Then, using the presence and absence of species across sites as the response variable, we sampled the posterior distributions of the baseline model, accounting for the information in the floristic database regarding species' indicator values and range of variation. Studying the relationship between the mean and variance of the distributions, we found clear evidence of the Rapopor's rule. Specifically, we found that species' range sizes decline with elevation and temperature (i.e.~$\beta_{i}$ and $\gamma_{i}$ in the baseline model are negatively correlated; Fig.~\ref{fig:correlation}; Supplementary Fig.~XX). The relationship was also found when using instead ordinal data (Supplementary Fig.~XX), but this pattern was not present along the second axis of variation of the environmental variables (Supplementary Fig.~XX). The comparison between the other parameter estimates revealed additional, somewhat more expected, relationships. In particular, we found the amplitude of distributions to be positively and negatively correlated with the mean and the variance along the elevation gradient, respectively (Supplementary Fig.~XX).

@@

Maintaining the symmetry of species' distributions, we then allowed the kurtosis---or shape of the tails---of these to vary in different ways. To do so, we changed the response curve of our Bayesian model to follow a generalized error distribution (Eq.~\ref{eq:generalized-response}). A comparison of the WAIC values showed this non-linear regression to outperform the baseline model (Supplementary Fig.~XX). Studying the resulting posterior distributions, we found the parameter controlling for the kurtosis to be centred around $\nu_i \sim 2$, which corresponds to a distribution with longer tails than a Gaussian (Fig.~\ref{fig:kurtosis-skewness}). However, such parameter estimates displayed a lot of variation across species, which might indicate that the shape of the tails of the distribution is species-specific (Supplementary Fig.~XX). Similarly, using Eq.~(\ref{eq:skewed-response}), we next studied the skewness of species distributions. Based on the estimates for the WAIC values, this model outperformed the previous models (Supplementary Fig.~XX), which sheds light on the naturally skewed nature of species' distributions. Perhaps most importantly, studying the mean value of the skewness across species (i.e.~parameter $\hat{\lambda}$), we found that species' distributions generally present steeper declines towards higher elevations (Fig.~\ref{fig:kurtosis-skewness}).

\begin{figure}[ht]
  \centering
    \includegraphics[width=0.9\textwidth]{figures/kurto-skew}
    	  \vspace{0.1cm}
	   \caption{Average kurtosis and skewness of species' distributions. Calculated using the posterior distributions of parameters $\hat{\nu}$ and $\hat{\lambda}$ from the models (see Supplementary Table XX), the two panels describe the average (a) kurtosis and (b) skewness of distributions, respectively. Panel (a) displays the results obtained by using a response curve that follows a generalized error distribution. Panel (b) displays the results obtained by using a response curve that follows a skewed normal distribution. In both case, the red dotted line indicates the conditions by which species are normally distributed along the environmental axis.}
      \label{fig:kurtosis-skewness}
      %7.5-3
\end{figure}

Finally, we used a model that allowed for heavy-tails and skewed response curves (Eq.~\ref{eq:heavy-skewed-response}) to understand the relationship between the kurtosis and skewness of distributions. This model outperformed the rest, presenting Akaike weights close to 1 when comparing models (Supplementary Fig.~XX). Comparing the posterior distribution for parameters $\nu_i$ and $\lambda_i$, we revealed the general constrains on the distribution of alpine plants studied. That is, we uncovered the relationship between the mean, variance, amplitude, kurtosis and skewness of empirical distribution of plants along an elevational gradient (Figs.~XX and YY). Moreover, different types of species seem to present characteristically different distributions (Fig.~XX). For example, invasive species... Notice the nature of these results does not depend on the presence or absence of a species at the edge of the sampling area (bootstrapping... Supplementary Fig.~XX). Likewise, the posterior distributions obtained using binary data were strikingly similar to those obtained using categorical data. This is important because...  is model identified similar relationships between parameters as the baseline model (Supplementary Fig.~XX), and measured similar average levels of kurtosis and skewness than the the other two models (Fig.~\ref{fig:kurtosis-skewness}).

%, and it does not seem to depend on species' type or reflect species' abundance change tendency over the years (Fig.~\ref{fig:universality}). T
%Moreover, species' physiology seems to strongly shape this parameter, which suggest that distribution skewness is an intrinsic property of species (New Fig.~XX).

\begin{figure}[ht]
  \centering
    \includegraphics[width=0.9\textwidth]{figures/means-Rapopor}
    	  \vspace{0.1cm}
	   \caption{Universality of the relationship between mean and variance of species' distributions. Comparison between how different types of species are mapped in Fig.~\ref{fig:correlation}a. Panel (a) describes the correlation coefficient between $\beta_i$ and $\gamma_i$ for each type of species. Panel (b) shows the differences between neophytes and native species in the way these are distributed along the environmental gradient. Panel (c) shows the same differences for species that have decreased, decreased in low elevations, increase and remain stable over the last decades (see Supplementary Table 1 for further details). {\color{red} I might move this to the Supplementary Information.}}
	   %7.5-4
      \label{fig:universality}
\end{figure}

@@

First, we test for the Rapopor's rule, which predicts a positive relationship between range size and elevation \citep{stevensElevationalGradientAltitudinal1992}. While this pattern has been largely studied for multiple systems and across gradients \citep{mccainElevationalRapoportRule2013}; contrasting evidence suggests this rule not to be pervasive across species \citep{ribasRapoportEffectWidespread2006, bhattaraiCanRapoportRule2006, mccainElevationalRapoportRule2013}. Our results not only allow us to properly test the existence of this geographical pattern, but they also showcase variation in how different types of species, such as native or neophytes, might respond to an environmental gradient. Second, we study whether or not species' distributions show steeper declines towards stressful conditions, testing the so-called abiotic stress limitation hypothesis (ref). \citet{normandImportanceAbioticStress2009} tested this for vegetation data using \citeauthor{huismanHierarchicalSetModels1993}'s statistical models for several independent species, finding no clear support for such a hypothesis. Our results are able to shed light on this geographical pattern as well as to highlight the degree to which different species will showcase different levels of decline towards stressful conditions. Specifically, we are able to link plant physiological traits to the skewness of their distributions.
